A continuous urbanization index to assess phylogenetic adaptations in urban bird living

### 2018-05-15 ms.Rmd
wording changes

# METHODS
## Study area
We looked at bird observations throughout the whole of continental Australia. Bird observations from outlying islands were eliminated from potential analyses.

## Bird observation data
We collated observations from eBird [@sullivan2009ebird; @sullivan2014ebird], a large-scale empirical dataset contributed by citizen scientists. Volunteer birdwatchers submit lists of species seen or heard at a given location, over a user-determined duration and survey area [@wood2011ebird]. We downloaded the eBird basic dataset (version ebd_relFeb-2018; available here: https://ebird.org/data/download), and used all observations between January 1, 2010 - February 28, 2018. This time-frame was used because this is the period in which eBird data are richest and because this time period minimized undue leverage of mismatch between changes in the response variable measurement (see below) over the given time frame in which bird observations were collected.

@@

## Phylogenetic analysis
For the phylogenetic portion of our analyses, we used the phylogenetic tree provided by Jetz et al. [@jetz2012global]. This phylogeny represents the first completely mapped phylogeny of 9,993 birds [@jetz2012global]. In order to assess whether there was a phylogenetic signal [@revell2008phylogenetic; @munkemuller2012measure] of the urbanization index and the continuous predictor variables used in our study, we used the phylosignal package [@keck2016phylosignal] in the R computing environment [@r2017coreteam]. This was also conducted for each of the continuous response variables in order to assess the relative contribution of phylogeny on the predictor variables.

@@

### Phylogenetically constrained models
Phylogenetic models were fit based on the same global model as above, but using the 'phylolm' function from the phylolm package [@ho2014phylolm]. Similar to above, we model-averaged the global phylogenetic model using the aforementioned methods.

## Taxonomic considerations
Given the diverse datasets used in this analysis, there were differing taxonomic authorities followed. The response data from eBird follows the eBird/clements checklist [http://www.birds.cornell.edu/clementschecklist/], while the biological and ecological data follows the working list of Australian birds by BirdLife Australia [http://www.birdlife.org.au/conservation/science/taxonomy], and the phylogenetic tree followed the BirdLife V3 world checklist [http://birdtree.org/taxonomy/]. There were minor differences which existed in common names, scientific names, and recognized species among the three datasets. A taxonomic key was made which combined the three different taxonomies (Appendix 2). Ultimately, a species' response variable, calculated from the eBird database and taxonomic source, was considered for analysis if it matched for both the trait dataset and the phylogenetic tree dataset. If it did not match both, then it was not considered for analysis (e.g., species currently only recognized by eBird/clements such as Western Whistler, Silver-backed Butcherbird, or Paperbark Flycatcher). Additionally, one species (Aleutian Tern) met the criteria for number of observations above, but was not included in the Australian trait data because it is new to Australia and thus not included in the analysis.

@@

# DISCUSSION
Niche breadth, as measured by different types of generalism, was significantly associated with our urbanization index across both phylogenetic and non-phylogenetic analyses, confirming the notion that urban birds have broad environmental tolerance [@bonier2007urban]. Conversely, granivores and insectivores were consistently negatively associated with urbanization. Furthermore, we found a significant phylogenetic signal [@revell2008phylogenetic] of the urbanization index in our analysis, suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation; @la2018phylogenetic]. This contrasts with previous studies [@evans2011makes], but this contrasting finding is likely a result of increased statistical power from our increased sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes]. Our results demonstrate that considering phylogeny in regards to the association between ecological and life history traits and urbanization in birds is important.

@@

Our finding that ecological generalists are more readily observed in urban areas confirms that of other work [@mckinney2006urbanization; @bonier2007urban; @moller2009successful]. We also confirm work which has found a relationship between annual fecundity and urbanization in birds [@croci2008does; @moller2009successful]. While some studies have found a lack of evidence that relative brain size is critical in urban birds [@kark2007living; @evans2011makes], we confirm those which do support the hypothesis that brain size is heavily correlated with urban birds [@moller2009successful; @maklakov2011brains; @moller2015brain]. Furthermore, we do confirm that urbanization appears to select against insectivorous species [@chace2006urban; @evans2011makes]. While others have found that urbanization supports granivorous species [@chace2006urban; @kark2007living; @evans2011makes], we found the opposite trend, but this could be a result of relatively little supplementary feeding in Australian cities [cf., @fuller2008garden]. We did not find any strong support for our hypotheses on different nesting types, but we did find that in general, species which tended to be more social were at an advantage to city living, confirming previous studies [@coleman1994neophobia; @jokimaki1998distribution; @kark2007living].

>>>>

# METHODS

## Bird observation data
We collated bird observations throughout the whole of continental Australia from eBird [@sullivan2009ebird; @sullivan2014ebird], a large-scale empirical dataset contributed by citizen scientists. Volunteer birdwatchers submit lists of species seen or heard at a given location, over a user-determined duration and survey area [@wood2011ebird]. We downloaded the eBird basic dataset (version ebd_relFeb-2018; available here: https://ebird.org/data/download), and used all observations between January 1, 2010 - February 28, 2018. This time-frame was used because this is the period in which eBird data are richest and because this time period minimized undue leverage of mismatch between changes in the response variable measurement (see below) over the given time frame in which bird observations were collected. Bird observations from outlying islands were eliminated from potential analyses.

@@

@@

### Phylogenetic analysis

For the phylogenetic portion of our analyses, we used the phylogenetic trees provided by Jetz et al. [@jetz2012global]. This phylogeny represents the first completely mapped phylogeny of 9,993 birds [@jetz2012global]. In order to assess whether there was a phylogenetic signal [@revell2008phylogenetic; @munkemuller2012measure] of the urbanization index and the continuous predictor variables used in our study, we used the phylosignal package [@keck2016phylosignal] in the R computing environment [@r2017coreteam]. This was also conducted for each of the continuous response variables in order to assess the relative contribution of phylogeny on the predictor variables.

Phylogenetic models were fit based on the same global model as above, but using the 'phylolm' function from the phylolm package [@ho2014phylolm]. Similar to above, we model-averaged the global phylogenetic model using the aforementioned methods.  We also  fit the same model to a range of the equally likely trees from [@jetz2012global] examining the distribution of putatively important parameters across the distribution of trees.

## Taxonomic considerations
Given the diverse datasets used in this analysis, there were differing taxonomic authorities followed. The response data from eBird follows the eBird/clements checklist [http://www.birds.cornell.edu/clementschecklist/], while the biological and ecological data follows the working list of Australian birds by BirdLife Australia [http://www.birdlife.org.au/conservation/science/taxonomy], and the phylogenetic tree followed the BirdLife V3 world checklist [http://birdtree.org/taxonomy/]. There were minor differences which existed in common names, scientific names, and recognized species among the three datasets. A taxonomic look-up table was made which combined the three different taxonomies (Appendix 2). Ultimately, a species' response variable, calculated from the eBird database and taxonomic source, was considered for analysis if it matched for both the trait dataset and the phylogenetic tree dataset. If it did not match both, then it was not considered for analysis (e.g., species currently only recognized by eBird/clements such as Western Whistler, Silver-backed Butcherbird, or Paperbark Flycatcher). Additionally, one species (Aleutian Tern) met the criteria for number of observations above, but was not included in the Australian trait data because it is new to Australia and thus not included in the analysis.

@@

# DISCUSSION

Niche breadth, as measured by different types of generalism, was significantly associated with our urbanization index.  This result was robust across non-phylogenetic and phylogenetic models, model averaging approaches, and across phylogenetic uncertaintly. Conversely, specilized feeding strategies were negatively associated with the urbanization index: granivore and insectivore specialists were much less likely to be found in urban environments. These results suggest that urbanization is, in general, excluding most specialist species.  SOMETHING ABOUT IMPLICATIONS FOR CONSERVATION / FRAGMENTATION.

The non-random distribution of the urbanization index across the phylogeny (Table 1) is interesting.  This result contrasts with previous studies [@evans2011makes], but this contrasting finding is likely a result of increased statistical power from our larger sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes].  Interestingly many of the predictor traits---generalism, clutch size---have even stronger phylogenetic signal, so it's not surprising that a larger sample size finds a non-random pattern in urbanization across the bird phylogeny.  It is worth noting that this results is consistent with the observation of reduced phylogenetic diversity in urban areas [@la2018phylogenetic], suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation; @major2010museum].

@@

Brain size at a given body size is an especially interesting trait--it's becomes more important in the phylogenetic model compared to the non-phylogenetic model.  While some studies have found a lack of evidence that relative brain size is critical in urban birds [@kark2007living; @evans2011makes], we confirm those which do support the hypothesis that brain size is heavily correlated with urban birds [@moller2009successful; @maklakov2011brains; @moller2015brain]. Because of the phylogenetic signal in both brain size and urbanization, it is crutial to investigate this relationship in a phylogenetically informed manner.

Certian aspects of the trait-urbanization relationship appear to consistent across the globe. CHANGE PHRASING BELOW HERE: The relationship with annual fecundity and urbanization in birds is consistent with work in Europe  [@croci2008does; @moller2009successful]. The that urbanization appears to select against insectivorous species [@chace2006urban; @evans2011makes]. While others have found that urbanization supports granivorous species [@chace2006urban; @kark2007living; @evans2011makes], we found the opposite trend, but this could be a result of relatively little supplementary feeding in Australian cities [cf., @fuller2008garden]. We did not find any strong support for our hypotheses on different nesting types, but we did find that in general, species which tended to be more social were at an advantage to city living, confirming previous studies [@coleman1994neophobia; @jokimaki1998distribution; @kark2007living].
